---
title: "Introduction to CovidWasteWatch"
author: "Vivian Wang"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: false
vignette: >
  %\VignetteIndexEntry{A tour of CovidWasteWatch}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(knitr)
opts_chunk$set(fig.align = "center", 
               out.width = "90%",
               fig.width = 6, fig.height = 5.5,
               dev.args=list(pointsize=10),
               par = TRUE, # needed for setting hook 
               collapse = TRUE, # collapse input & ouput code in chunks
               warning = FALSE)
library(CovidWasteWatch)
```


## Introduction

`CovidWasteWatch` is an R package developed to provide analysis and visualization 
options for COVID-19 wastewater viral signal data and viral variant frequency data.

To download **CovidWasteWatch**, use the following commands:

``` r
require("devtools")
devtools::install_github("viv-wang/CovidWasteWatch", build_vignettes = TRUE)
library("CovidWasteWatch")
```
To list all sample functions available in the package:
``` r
ls("package:CovidWasteWatch")
```

To list all sample datasets available in the package:
``` r
data(package = "CovidWasteWatch")
```

<br>

## Components

<div style="text-align:center"><img src="" alt="" width="500"/>


## Applications

### Viral signal data

Provide a valid path to a CSV file which includes columns 'date' and 'signal'.
In this example, we use a dataset from COVID-19 wastewater surveillance by Public
Health Ontario (see data documentation for more details).

```{r}
# Get path to example viral signal data CSV file
signalFilepath <- system.file("extdata", "viral_signal_input_data.csv", 
                              package = "CovidWasteWatch")
# Run and store results of ViralSignal()
ViralSignalResults <- ViralSignal(fileSignal = signalFilepath)

# To view each component of the result separtely, see names 
names(ViralSignalResults)
# Alternatively, view all results
ViralSignalResults
```

The plotting function `PlotViralSignal` can also be called on separately,
with a dataframe that follows the required format.
Here we use the dataframe containing data from the example dataset that has 
been processed `ViralSignal()`: `ViralSignalDF` (see data documentation for more details).

```{r}
data("ViralSignalDF")
SignalPlot <- PlotViralSignal(signalData = ViralSignalDF)
# View plot
SignalPlot
```


### Viral variant frequency (proportion) data

Provide a valid path to a CSV file which includes columns 'date', 'variant', 
'proportion', and, optionally 'parent' for parent variants. If a 'parent'
column is not included, set the parameter `parentVariant = FALSE`.

Here we use a public dataset of weekly COVID-19 variant breakdown from Health Canada
(see data documentation for more details). This file includes a 'parent' column.

```{r}
# Get path to example variant breakdown CSV file
varFilepath <- system.file("extdata", "variant_input_data.csv", package = "CovidWasteWatch")
# Run and store results of VarBreakdown()
# Note: parentVariants = TRUE by default
VarBreakdownResults <- VarBreakdown(fileVariant = varFilepath)

# To view each component of the result separtely, see names 
names(VarBreakdownResults)
# Alternatively, view all results
VarBreakdownResults
```

The plotting function `PlotVarBreakdown` can also be called on separately,
with a dataframe that follows the required format.
Here we use the dataframe containing data from the example dataset that has 
been processed `VarBreakdown()`: `VariantDF` (see data documentation for more details).

```{r}
data("VariantDF")
VarPlot <- PlotVarBreakdown(varData = VariantDF)
# View plot
VarPlot
```

<br>

## Package References

- Wang, V. (2023) CovidWasteWatch: An R Package for COVID-19 Wastewater Data 
  Analysis and Visualization. Unpublished. https://github.com/viv-wang/CovidWasteWatch

<br>

## Other References

- Bengtsson, H (2022). R.utils: Various Programming Utilities. https://henrikbengtsson.github.io/R.utils/

- R Core Team (2022). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. https://www.R-project.org/

- Silva, A. (2022) TestingPackage: An Example R Package For BCB410H.
  Unpublished. URL https://github.com/anjalisilva/TestingPackage

- Wickham, H. (2016). ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York. ISBN 978-3-319-24277-4. https://ggplot2.tidyverse.org

- Wickham H, François R, Henry L, Müller K, Vaughan D. (2023). dplyr: A Grammar of Data Manipulation. https://dplyr.tidyverse.org

----

```{r}
sessionInfo()
```
